\LARGE{{\bf{Piled up gravitational waves:} \\ \bf{Searching for new signals of $N$naturalness} }}

### 2019-08-16 gwnnDraft02.tex
signals update

\documentclass[nofootinbib,twocolumn,preprintnumbers]{revtex4-2}
%\usepackage[dvips]{graphicx}
\pdfoutput=1
\usepackage{amsmath,amsthm,amssymb,multirow,psfrag}
\usepackage{epsfig}
\usepackage{color}
\usepackage{slashed}
\graphicspath{{./Figures/}}

@@

\section{Gravitational Waves}
\label{sec:gw}

\section{Particle Setup}
\label{sec:nn}

@@

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{NeffBounds.png}
\label{fig:Neff}
\hfill
\caption{$\Delta N_{eff}$ as a function of relativistic degrees of freedom in a fully decoupled hidden sector ($g_h$) and the temperature ratio of said sector. Contours set $\Delta N_{eff}$ to 0.01 (forest green), 0.03 (lime green), 0.1 (gentle pink), and 0.5 (royal violet). Most stringent current $N_{eff}$ bounds shade in grey. 
}
\end{figure}

@@

\section{Signals}
\label{sec:signals}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{apsrev4-2}
\bibliography{refs}

>>>>

\documentclass[nofootinbib,twocolumn,preprintnumbers]{revtex4-1}
%\usepackage[dvips]{graphicx}
\pdfoutput=1
\usepackage{amsmath,amsthm,amssymb,multirow,psfrag}
\usepackage{epsfig}
\usepackage{color}
\usepackage{slashed}
\graphicspath{{./Figures/}}

@@

\section{Gravitational Waves}
\label{sec:gw}
Sectors other than the SM can be reheated through multifield inflationary theories, Reheaton models, etc. In these models, the SM will receive the majority of the energy density while the remaining energy density will be distributed, in a model dependent way, to the various hidden sectors. If there does not exist a portal between the sectors, they will remain decoupled and evolve independently. When looking at multiple thermally decoupled hidden sectors it is convienent to define the temperature ratio [cite cold,dark,noisy]
\begin{equation}
\xi_i \equiv  \frac{T_{i}}{T_{\gamma}}
\end{equation}
Where $T_{i}$ is the temperature of the $i^{th}$ hidden sector, and $T_{\gamma}$ the temperature of the standard model (assuming radiation domination).

\vspace{1em }
- first order phase transitions ( strength, scale, origin)
Right after reheating, the temperatures of the respective sectors are high leading to restoration of a given sector's symmetry.  An example being the SM's $SU(3) \times SU(2) \times U(1)$ symmetry being restored as the scalar potential's vev becomes global minimum. As the sectors cool down their scalar potentials may change, producing a new global vev at which the symmetry will break by an induced phase transition. Depending on the characteristics of the symmetry breaking, the phase transition may be strongly first ordered (SFOPT).

Once the temperature reaches the critical temperature the SFOPT will occur, at which bubbles of the new phase will start to nucleate. These bubbles will expand, eventually colliding and merging until the entire sector is within the new phase.  These bubbles are discribed by the following Euclidean action

\begin{equation}
S_{E}(T) = \frac{1}{T}\int d^3x \bigg[\frac{1}{2}(\nabla\phi)^2 + V(\phi,T)  \bigg]
\end{equation}

Where the time component has been integrated out due to finite temperature effects. The system has solutions with a $O(3)$ symmetry. We leave the thermalized potential $V(\phi, T)$ general, since an exact QCD potential at the time of the chiral phase transition is not well understood. Some authors have used the chiral effective lagrangian to calculate a low energy thermalized potential for confining $SU(N)$ .

The amount of energy density dumped into the individual sectors dictates the energy budget for the PT and hence for the gravitational waves.  A quantity that characterizes the strength of the PT is the ratio of the latent heat and energy density of radiation, at the time of nucleation, 
\begin{equation}
\alpha = \frac{\epsilon}{g_{*} \pi^2 (T^{nuc}_{\gamma})^4/30}
\end{equation}
Assuming that there is a negligible amount of energy being dumped back into the SM radiation bath, which would cause significant reheating of $\rho_{\gamma}$, the latent heat of the PT $\epsilon$ should  approximately be the energy density of the hidden sector going through the PT. $g_{*}$ has weak temperature dependence in a single sector, but when dealing with multiple hidden sectors, $g_{*}$ gains contributions from all $N$ sector's relativistic degrees of freedom, weighted by their respective temperature ratios (energy densities).
\begin{equation}
g_{*} = g_{*,\gamma} + \sum_{i = 1}^{N} g_{*,i} (\xi_{i})^4
\end{equation}
It will be seen that for models with a significant number of hidden sectors, that the lack of substantial reheating in adjecent sectors will cause their respective $g_{*}$ corrections to become negligible, $g_{*} \approx g_{*,\gamma}$.
\section{Particle Setup}
\label{sec:nn}

@@

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{neff.png}
\label{fig:Neff}
\hfill
\caption{$\Delta N_{eff}$ as a function of relativistic degrees of freedom in a fully decoupled hidden sector ($g_h$) and the temperature ratio of said sector. Contours set $\Delta N_{eff}$ to (0.01, 0.03,  0.1,  0.5) as the dashed lines, from left to right. Most stringent current $N_{eff}$ bounds shade in purple. 
}
\end{figure}

@@

\section{Signals}
\label{sec:signals}
\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=.45\textwidth]{TwoFar.png} 
\hfill
\includegraphics[width=.45\textwidth]{TwoMed.png} 
\hfill
\includegraphics[width=.45\textwidth ]{highest.png}
\hfill
\includegraphics[width=.45\textwidth]{energydensity.png} 
\end{minipage}
\hfill
\caption{ Gravitational wave power spectrum for various models. The top row consists of two models with only two hidden sectors undergoing SFOPT at differing temperatures. The bottom right (left) are examples of exotic sectors, discussed in section V, with sectors reheated at $\rho \sim 0.25\rho_{SM}$ ($\rho \sim 0.17\rho_{SM}$ ). All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$, with $\beta/H  = 10$.  The shaded curves are the power law noise curves calculated from expected sensitivity curves for space-based interferometers and pulsar timing arrays; Lisa ( blue), DECIGO (light blue), BBO (red), SKA 5-year (purple), SKA 10-year (orange), SKA 20-year (green)   }
\label{fig:Haa}
\end{figure*}
Gravitational waves are produced with contributions from different components of the SFOPT's evolution.  The three leading order contributions to the GW power spectrum are as follows:
\begin{itemize}
\item \textbf{Scalar field contributions $\Omega_{\phi}$}: Caused by collisions of the bubble walls, the solutions being completely dependent on the scalar field configuration. With efficiency factor $\kappa_{\phi} = 1 - \alpha_{\infty}/\alpha$.
\item \textbf{Sound wave contributions $\Omega_{v}$}: Sound waves within the plasma after bubble collision will produce $\beta/H$ enhanced gravitational waves. With efficiency factor $\kappa_{v} \propto \alpha_{\infty}/\alpha$.
\item \textbf{Magnetohydrodynamics contributions $\Omega_{B}$}: MTG shocks within the plasma, left over from the sound wave propagation, will produce gravitational waves. With efficiency factor $\kappa_{
turb} \approx 0.1 \kappa_{v} $.
\end{itemize}
The total gravitational wave power spectrum is a linear combination of the three contributions,
\begin{equation}
h^2\Omega_{GW} \approx h^2\Omega_{\phi} + h^2\Omega_{v} + h^2\Omega_{turb} 
\end{equation}
In essence, this is a linear combination of the individual spectral shapes weighted by their efficiency factors $\kappa$. As shown in section IV, the mass spectrum from the dark QCD chiral breaking sectors is such that the runaway bubble condition is satisfied ($\alpha >\alpha_{\infty}$). In this case, the efficiency factors for sound wave and MHD contributions tend to zero and the GWs produced are purely from bubble collisions, $h^2\Omega_{GW} \approx h^2\Omega_{\phi}$.  The form of the GW power spectrum at the time of nucleation is given as,

\begin{equation}
h^2\Omega_{GW} = 7.7\times 10^{-2} \bigg( \frac{\kappa_{\phi} \alpha}{1 + \alpha} \bigg)^2 \bigg( \frac{H}{\beta} \bigg)^2  S(f)
\end{equation}
Where we have assumed an optimistic bubble wall velocity of $v \sim c$ for runaway bubbles. $S(f)$ is the  frequency spectrum, and a parametric from has been found through numerical simulations of bubble wall collisions.
\begin{equation}
S(f) = \frac{3.8 \,(f/f_p)^{2.8}}{1 + 2.8\,(f/f_p)^{3.8}}
\end{equation}
The peak frequency, $f_{p}$ is a function of the temperature of the SM at the time of nucleation. Depending on the model, the various hidden sectors can phase transition at different scales and therefore temperatures, cause a shift in the GW spectrum's peak frequency given by,

\begin{equation}
f_{p} = 3.8 \times 10^{-8} \, \textrm{Hz}\, \bigg( \frac{\beta}{H}\bigg)\bigg(\frac{T_{\gamma}}{100 \, \textrm{GeV}}\bigg)\bigg(\frac{g_{*}}{100}\bigg)^{\frac{1}{6}}
\end{equation}

Now that the framework has been laid out for the creation of GW from a single SFOPT, we generalize to multiple sectors going under independent, coherent SFOPT. In the models presented in this paper, we consider a subset of N-hidden sectors that phase transition at a SM temperature of $T_{\gamma}^{i}$. As the GWs propagate in free space, the energy density and frequency spectrum, at the time of production $\Omega_{GW}^{*}(f)$, will redshift to today's value $\Omega_{GW}^{0}(f) = \mathcal{A}\, \Omega_{GW}^{*}((a_{0}/a) f) $. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW power spectrum that would be measured today is given by the coherent sum,

\begin{equation}
\Omega_{GW}^{0} =  \sum_{i}^{N} \mathcal{A}^{i}\, \Omega_{GW}^{i,*}((a_{0}/a)_{i} f) 
\end{equation}

It is safe to assume that a subset of parameters of the SFOPT do not differ between sectors within Nnatrualness or other decoupled hidden sector models. In the case of N-copies of dark QCD-like phase transitions, the relativistic degrees of freedom, phase transition rate,  the dark QCD scale, are all similar. This makes the redshifting factor, $\mathcal{A}^i$, independent of sector number. We can also see that the frequency dependence in the spectrum takes the form of $f/f_{p}$, this causes a cancellation of the redshifting factors in the ratios. As multiple sectors phase transition at different times, and therefore temperatures, the peaks will shift relative to each other, purely from the linear temperature dependence of the peak frequency $f_{p} \sim T_{\gamma}$.  This is seen in figures [], where the spectrum peaks are shifted causing a peak broadening of the convoluted spectrum. The broadening can be substantial if the hidden sectors transition between a large gap of time (temperature). Eventually, a temperature limit will be reached at which two (or multiple) peaks, if the amplitudes are comparable.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{apsrev4-1}
\bibliography{refs}
